---
title: "DSO 522 Group 5 Airbnb Pricing Analysis"
author: "Group 5 (Evelyn (Wanyi) Dong, Qiongqiong Lin, Jenny Shang, Yoki (Lingrou) Wang) "
date: "11/1/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: \usepackage{color}
---

\begin{center} 
{\color{red} \textbf{DSO 522 Group 5 Airbnb Pricing Analysis}}
\end{center}

For this project, we want to study the impact of Covid-19 on the tourism industry by
analyzing past price trends of Airbnb rental properties in the city of Oakland, California, and predict future rental prices by applying different time-series forecasting methods. We choose this topic because Airbnb rentals have seen significant growth in the past but are now struggling from booking declines due to Covid-19. We are interested in using time-series forecasting to predict future prices and provide valuable insights and recommendations to help Airbnb’s stakeholders develop a solid pricing strategy to achieve customer retention and sustainable growth. From a technical standpoint, Airbnb data is readily available on Airbnb’s official website (http://insideairbnb.com/get-the-data.html). It is pre-processed, well-structured and standardized, making it the ideal dataset to conduct our analysis.

The complete dataset represents 722 Oakland average daily Airbnb house pricing from July 16, 2018 to July 16, 2020. 

## Our analysis is built on three key objectives: 
(1.) Understand the historical patterns of the Airbnb data by studying potential trends and  periodic components. 
```{r}
# import packages
library(zoo)
```

```{r}
library(forecast)
# read data
data = read.csv("cleaned_data.csv")
# set zoo object of price
price = zoo(data$price, seq(from = as.Date("2018-07-16"), to = as.Date("2020-07-16"), by = 1))
# plot the data
autoplot(price, xlab=("Time") , ylab=("Average listing price") )
```

```{r Loading Data in TS format}
library(forecast)

price.ts = ts(data$price, start = c(2018,197), frequency=365)
ma.trailing.roll.week = rollmean(price.ts, k = 7, align = "right")
ma.trailing.roll.month = rollmean(price.ts, k = 30, align = "right")


# plot the data
autoplot(price.ts,main='Airbnb Oakland Daily Pricing', 
         xlab=("Time") , ylab=("Average listing price") )+
  autolayer(ma.trailing.roll.week,series='Weekly Aggregated')+
  autolayer(ma.trailing.roll.month,series='Monthly Aggregated')
  
```

1. Ask for advice on predicting daily v.s. monthly price of Oakland (monthly with very limited data points but does have monthly alternative data available)
Depends on the scope of what we wanted to use them for.
Using alternative data that's not on the same level of details - extrapolation issue - check to see what are their performances.


** Look into why 2019 sharp drop
** Qualitative external variable : use dummy variable to indicate those exceptional cases?
2. Ask for clarification on COVID impact on model buildling - seperate models for post COVID?
By looking at the data we seee a sharper decrease in 2019 compare with COVID's impact:
* Build the model without 2020, and then make predictions in 2020



```{r Smoothing}
changeofprice = read.csv("Change_of_Price.csv")
changeofprice <- ts(changeofprice$X..Change,start=c(2018,198),frequency=365)
autoplot(changeofprice,main='Change of Price Lag 1',ylab='% of Change in Price')

library(lmtest)
#AR(1)
ar1<-Arima(changeofprice,order=c(1,0,0))
coeftest(ar1)

```

We see that the coefficient is far away from 0 and P value also did not show statistical significance, which means this is not a random walk.






Key findings 
(-) The pricing data shows more variability in 2018 and 2020 than 2019. 
(-) In 2018, the price shows a donward trend in the beginning, but resumes in November and reaches a peak in December (potential reason: Thanksgiving, Christmas and New Year).
(-) In 2019, the listing price shows an upward trend. 
(-) In 2020, the listing price drops in the beginning (potential reason: covid-19); however, starting from May, the price resumes and increases. 
(-) Weekly seasonality is shown in the data. Since we only have two years data, it is hard to tell if there is a yearly seasonality. 

(2.) Explore a variety of time-series models and determine the most appropriate candidate  model on the basis of accuracy and simplicity to analyze and forecast future prices. 

(-) 
```{r}

```

(3.) Implement the selected candidate model to forecast both short-term and long-term Airbnb  average listing price in Oakland, California. 

## With our objectives in mind, we strive to deliver to Airbnb stakeholders a guidance on listing  price. Our analysis will focus on answering four important questions: 

(1.) What are the  determinative factors of average listing prices? 

(2.) Are there any lags between Covid-19 outbreak  and rental prices? 

(3.) Have vacation rentals recovered from the pandemic? 

(4.) Are there any  important external factors that play a role in determining the average listing price? In the future, we can apply our analysis to broader geographic areas. 

Some important variables in our study include the overall average listing price of the  selected city, and the average listing price per room type (entire home/apt or private room).  Additional variables we want to consider include the city’s unemployment rate over time, which  can be found on the U.S. Bureau of Labor Statistics website (https://www.bls.gov), as well as  general economic indicators such as GDP growth, which can be obtained from St. Louis Fed  (https://fred.stlouisfed.org/). 

At the end of our study, we expect to draw critical insights from historical patterns in  order to assist Airbnb’s stakeholders. We aim to provide Airbnb hosts with an understanding of  the determinative factors affecting property listing price such that they can adjust ahead of time,  suggest an accurate, simple and effective time-series forecasting model for Airbnb to predict  average short-term rental pricing and long-term rental pricing, and explore potential solutions for  Airbnb to identify risks and launch effective promotion or recovery plans in this Covid-19 era.

