---
title: "DSO 522 Group 5 Airbnb Pricing Analysis"
author: "Group 5 (Evelyn (Wanyi) Dong, Qiongqiong Lin, Jenny Shang, Yoki (Lingrou) Wang) "
date: "11/1/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: \usepackage{color}
editor_options: 
  chunk_output_type: console
---

\begin{center} 
{\color{red} \textbf{DSO 522 Group 5 Airbnb Pricing Analysis}}
\end{center}

For this project, we want to study the impact of Covid-19 on the tourism industry by
analyzing past price trends of Airbnb rental properties in the city of Oakland, California, and predict future rental prices by applying different time-series forecasting methods. We choose this topic because Airbnb rentals have seen significant growth in the past but are now struggling from booking declines due to Covid-19. We are interested in using time-series forecasting to predict future prices and provide valuable insights and recommendations to help Airbnb’s stakeholders develop a solid pricing strategy to achieve customer retention and sustainable growth. From a technical standpoint, Airbnb data is readily available on Airbnb’s official website (http://insideairbnb.com/get-the-data.html). It is pre-processed, well-structured and standardized, making it the ideal dataset to conduct our analysis.

The complete dataset represents 722 Oakland average daily Airbnb house pricing from July 16, 2018 to July 16, 2020. 

## Our analysis is built on three key objectives: 
(1.) Understand the historical patterns of the Airbnb data by studying potential trends and  periodic components. 

```{r}
# import packages
library(zoo)
library(forecast)
library(ggplot2)
library(dplyr)
```

```{r}
# read data
data = read.csv("cleaned_data.csv")
# set zoo object of price
price = zoo(data$price, seq(from = as.Date("2018-07-16"), to = as.Date("2020-07-16"), by = 1))
# plot the data
autoplot(price, xlab=("Time") , ylab=("Average listing price") )
# load data in ts format 
## price.ts = ts(data$price, start = c(2018,197), frequency=7)
price.ts = ts(data$price, start = c(2018,197), frequency=365)
# function 1 
price.1 = window(price.ts, end=c(2020, 13))
nValid = 60
price.1.train = window(price.1, end=c(2020, 13-nValid))
price.1.test = window(price.1, start=c(2020, 13-nValid+1))
# function 2
price.2.train = window(price.ts, end=c(2020, 188-nValid))
price.2.test = window(price.ts, start=c(2020, 188-nValid+1))
```



```{r Functin1}
## Set validation days
numValid = 30

## Create empty vectors to store results
naive_pred = rep(NA,60)
snaive_mape_train = rep(NA,numValid)
snaive_mape_test = rep(NA, numValid)
```


```{r}
for(i in 1:numValid){

  # Split the data into training and validation

  nTrain = length(price.1.train) + i
  train.ts = window(price.1, start = c(2018,197), end = c(2018, 197+nTrain-1))

  # Fit the Seasonal Naive Model
  naive_1step<-snaive(train.ts,h=numValid)

  # Make prediction
  pred = forecast(naive_1step, h=numValid)

  # Store the MAPE (Testing MAPE is the 10th value)
  snaive_mape_train[i] = accuracy(naive_1step)[5]
  snaive_mape_test[i] = accuracy(pred$mean,price.1.test[i:(i+numValid-1)])[5]
  
}

# Average of the training and testing MAPE is the error of the model
mean(snaive_mape_train)
mean(snaive_mape_test)
```


```{r Functin2}
## Set validation days
numValid = 30

## Create empty vectors to store results
naive_pred = rep(NA,60)
snaive_mape_train = rep(NA,numValid)
snaive_mape_test = rep(NA, numValid)
```

```{r}
for(i in 1:numValid){

  # Split the data into training and validation

  nTrain = length(price.2.train) + i
  train.ts = window(price.ts, start = c(2018,197), end = c(2018, 197+nTrain-1))

  # Fit the Seasonal Naive Model
  naive_1step<-snaive(train.ts,h=numValid)

  # Make prediction
  pred = forecast(naive_1step, h=numValid)

  # Store the MAPE (Testing MAPE is the 10th value)
  snaive_mape_train[i] = accuracy(naive_1step)[5]
  snaive_mape_test[i] = accuracy(pred$mean,price.2.test[i:(i+numValid-1)])[5]
  
}

# Average of the training and testing MAPE is the error of the model
mean(snaive_mape_train)
mean(snaive_mape_test)
```

